---
title: "lclinton_hmk5"
author: "Laurence Clinton"
date: "February 8, 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1.a First, import the .txt file into R so you can process it.



```{r ReadBabyFile}
#set working directory
setwd("D:/SMU/DoingDataScience/Unit5")

#read in 'yob2016' file
BabyFile <- read.delim("yob2016.txt", header = FALSE,sep=";")
head(BabyFile)

# create data frame from read file
df <- data.frame(BabyFile)
names(df) <- c("First Names","Gender","Name Count")

# display the first few rows
head(df)
```

## 1.b Display the summary and structure of df

Output from: summary(df) and str(df)

```{r SummaryStructure}
#summary of data frame
summary(df)
#structure of data frame
 str(df)

```
## 1.c Write an R command to figure out which name it is and display it.

```{r FindOddName}
# grep names that end in 'yyy'
grep("yyy$",df$`First Names`)
# confirm finding
df[212,]

```
## 1.d Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant. Save the remaining dataset as an object: y2016
```{r RemoveName}
# Remove observation and set to object y2016
y2016 <- df[-c(212),]

#Compare before and after
nrow(df)

nrow(y2016)


```

## 2.a Like 1a, please import the .txt file into R. Look at the file before you do. You might have to change some options to import it properly. Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.

```{r ReadIn2015}
#read in file
BabyFile2015 <- read.delim("yob2015.txt", header = FALSE,sep=",")
#display first few rows
head(BabyFile2015)

# create data frame from file
y2015 <- data.frame(BabyFile2015)
names(y2015) <- c("First Names","Gender","Name Count")

```
## 2.b Display the last ten rows in the dataframe. Describe something you find interesting
about these 10 rows

```{r lastRows2015}
# show last 10 rows
tail(y2015,10)

#There are some unsual but similar names that start with 'Zy'.

```
## 2.c Merge y2016 and y2015 by your Name column; assign it to final. The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.

```{r merge}
#merge data from y2015 and y2016
final2 <- merge(y2015,y2016,by="First Names")
#show top rows
head(final2)

```
## 3.a  Create a new column called "Total" in final that adds the amount of children in 2015 and 2016 together. In those two years combined, how many people were given popular names.


```{r Total}
final2$Totals <- final2$`Name Count.x` + final2$`Name Count.y`

#How many were given popular names?

#The top ten names seem popular and total 343,758.

```
## 3.b  Sort data by Total.  What were the top 10 most popular names?
```{r Final}
final2_order <- final2[order(final2$Totals,decreasing = TRUE),]


#Display top 10 rows
head(final2_order,10)

```

## 3.c Omit boys.  Top 10 most popular girl's names

```{r Omit boys}
head(final2_order[final2_order$Gender.x == "F",],10)

```

## 3.d  Write top 10 girl names and totals to csv.
```{r write to csv}
#set final data frame
final_f <- head(final2_order[final2_order$Gender.x == "F",],10)
final_df <- data.frame(final_f$`First Names`,final_f$Totals)

#Update column names
names(final_df) <- c("First Names","Totals")

#Display rows
head(final_df,10)

#write to csv
write.csv(final_df, file="Female_Names2.csv",row.names = FALSE)
```

